---
title: "Initial analyses"
author: "Jacob J. Bukoski"
date: "November 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}

library(leaflet)
library(lme4)
library(mapview)
library(raster)
library(sf)
library(sp)
library(spdep)
library(tidyverse)

```


```{r echo = FALSE, message = FALSE, warning = FALSE}

dat <- read_csv("/home/jbukoski/research/data/thailand_stocks/output/processed_subplot_data.csv")
dat_sf <- read_sf(dsn = "/home/jbukoski/research/data/thailand_stocks/output/", 
               layer = "processed_subplot_data.shp")

krabi <- dat %>%
  filter(site == "Krabi") %>%
  drop_na()

nakorn <- dat %>%
  filter(site == "Nakorn") %>%
  drop_na()

```

### Structure of datasets:

- Krabi
    - 7 transects
        - 5 plots per transect, all with aboveground characteristics
        - 4 plots per transect with soil and aboveground characteristics

- Nakorn
    - 10 transects
        - 5 plots per transect, all with aboveground characteristics
        - 2 plots per transect with soil and aboveground characteristics
            - Will collect additional 2 plots
      
### Locations of the data

Basic visualization of where the data sits. Data are summarized at the plot level, and attached to the markers on the map.

  - Can view plot-level data by clicking.
  - Can also adjust basemap layer.

```{r echo = FALSE}

mapview(dat_sf, map.types = c("Esri.WorldImagery", "OpenStreetMap", "Esri.WorldTopoMap"))

```

## Initial passes and visualizations of the data

#### Total biomass ~ Basal area @ plot level

```{r, echo = FALSE}

dat %>%
  ggplot() +
  geom_point(aes(x = subplot_ba, y = subplot_ttl_ha, col = site)) +
  xlab("BA") +
  ylab("Total Biomass (Mg/ha)")

```

#### Total biomass ~ plot

Relatively minor differences in biomass across sites, though slightly higher in Nakorn.

```{r, echo = FALSE}

dat %>%
  ggplot() + 
  geom_boxplot(aes(x = plot_id, y = subplot_ttl_ha, col = site)) +
  xlab("Transect") +
  ylab("Total biomass (Mg/ha)")

```

## Soil parameters across both sites

Can begin to see the differences in soil parameters at site level here. Plotted as scatter plots as Nakorn only has two plots per site.

#### Bulk density ~ transect

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  ggplot() + 
  geom_point(aes(x = plot_id, y = subplot_bd, col = site)) +
  xlab("Transect") +
  ylab("Bulk density (g/cm^3)")

```

#### Percent organic carbon ~ transect

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  ggplot() + 
  geom_point(aes(x = plot_id, y = subplot_poc, col = site)) +
  xlab("Transect") +
  ylab("Percent organic C (%)")

```

#### Soil carbon density (mg/cm^3) ~ transect

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  ggplot() + 
  geom_point(aes(x = plot_id, y = subplot_c_dens, col = site)) +
  xlab("Transect") +
  ylab("Carbon density (mg/cm^3)")

```

#### Soil organic carbon to mean depth (Mg/ha) ~ transect

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  ggplot() + 
  geom_point(aes(x = plot_id, y = subplot_soc, col = site)) +
  xlab("Transect") +
  ylab("Soil organic C (Mg/ha)")

```



# Looking at things by dominant genus/species

Summary of counts of plots of different dominant genus (plots from both sites pooled). Rh, Br, etc., are genus abbreviations. "Rh" alone is a plot with 80% dominance of that genus, "Rh_mix" is plot dominated by that genus but < 80%.

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  group_by(dom_genus) %>% 
  add_count(dom_genus) %>% 
  select(dom_genus, n) %>% 
  distinct()

```

Structure of the dominant genus-labeled plots at the site level - pretty unbalanced. Rhizophora dominated plots (rh_mix) are by far most prevalent in Krabi.

```{r, echo = FALSE}
dat %>% 
  drop_na() %>% 
  group_by(site, dom_genus) %>% 
  add_count(dom_genus) %>% 
  select(site, dom_genus, n) %>% 
  distinct()

```

#### Total biomass ~ dominant genus

*Note*: Dominant genus is equally weighted importance index of frequency, dominance, and density at genus level. Dominant species is calculated in same fashion but at species level.

```{r, echo = FALSE}

dat %>%
  ggplot() + 
  geom_boxplot(aes(x = dom_genus, y = subplot_ttl_ha, col = site)) +
  xlab("Dominant genus") +
  ylab("Total biomass (Mg/ha)")

```

#### Bulk density ~ dominant genus

```{r, echo = FALSE}

dat %>% 
  drop_na() %>%
  ggplot() + 
  geom_boxplot(aes(x = dom_genus, y = subplot_bd, col = site)) +
  xlab("Dominant genus") +
  ylab("Bulk density (g/cm^3)")

```

#### Percent organic carbon (%) ~ dominant genus 

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  ggplot() + 
  geom_boxplot(aes(x = dom_genus, y = subplot_poc, col = site)) +
  xlab("Dominant genus") +
  ylab("Percent organic C (%)")

```

#### Soil carbon density (mg/cm^3) ~ dominant genus

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  ggplot() + 
  geom_boxplot(aes(x = dom_genus, y = subplot_c_dens, col = site)) +
  xlab("Dominant genus") +
  ylab("Soil C density (mg/cm^3)")

```

#### Soil carbon density (Mg/ha to mean soil depth) ~ dominant genus

```{r, echo = FALSE}

dat %>%
  drop_na() %>%
  ggplot() + 
  geom_boxplot(aes(x = dom_genus, y = subplot_soc, col = site)) +
  xlab("Dominant genus") +
  ylab("Soil organic carbon (Mg/ha)")

```

# Autocorrelation

## Autocorrelation at within site level for Krabi

#### Total biomass, Krabi

```{r, echo = FALSE}

krabi_sp <- dat_sf %>%
  drop_na() %>%
  filter(site == "Krabi") %>%
  as("Spatial")

nb <- knn2nb(knearneigh(krabi_sp, k = round(nrow(krabi_sp)*0.3, 0)))
nbm <- nb2mat(nb, style = "B")
nbw <- nb2listw(nb, style='B')

ttl_biom_mc <- moran.mc(krabi_sp$sbplt_t_, nbw, nsim=99)
ttl_biom_mc

plot(ttl_biom_mc, main = "Krabi Biomass spatial autocorrelation", xlab = NULL)

```


#### Soil bulk density, Krabi

```{r, echo = FALSE}

k_bd_mc <- moran.mc(krabi_sp$sbplt_bd, nbw, nsim=99)
k_bd_mc

plot(k_bd_mc, main = "Spatial autocorrelation: Krabi bulk density", xlab = NULL)

```

#### Soil percent organic carbon, Krabi

```{r, echo = FALSE}

k_poc_mc <- moran.mc(krabi_sp$sbplt_p, nbw, nsim=99)
k_poc_mc

plot(k_poc_mc, main = "Spatial autocorrelation: Krabi percent organic carbon", xlab = NULL)

```

#### Soil carbon density, Krabi

```{r, echo = FALSE}

k_cdens_mc <- moran.mc(krabi_sp$sbplt_c_, nbw, nsim=99)
k_cdens_mc

plot(k_cdens_mc, main = "Spatial autocorrelation: Krabi soil c density", xlab = NULL)

```



## Autocorrelation at within-site level for Nakorn

#### Total biomass, Nakorn

```{r, echo = FALSE}

nakorn_sp <- dat_sf %>%
  drop_na() %>%
  filter(site == "Nakorn") %>%
  as("Spatial")

nb <- knn2nb(knearneigh(nakorn_sp, k = round(nrow(nakorn_sp)*0.3, 0)))
nbm <- nb2mat(nb, style = "B")
nbw <- nb2listw(nb, style='B')

n_ttl_biom_mc <- moran.mc(nakorn_sp$sbplt_t_, nbw, nsim=99)
n_ttl_biom_mc

plot(n_ttl_biom_mc, main = "Spatial autocorrelation: Nakorn total biomass", xlab = NULL)

```

#### Soil bulk density, Nakorn

```{r, echo = FALSE}

n_bd_mc <- moran.mc(nakorn_sp$sbplt_bd, nbw, nsim=99)
n_bd_mc

plot(n_bd_mc, main = "Spatial autocorrelation: Nakorn soil bulk density", xlab = NULL)

```

#### Soil percent organic carbon, Nakorn

```{r, echo = FALSE}

n_poc_mc <- moran.mc(nakorn_sp$sbplt_p, nbw, nsim=99)
n_poc_mc

plot(n_poc_mc, main = "Spatial autocorrelation: Nakorn percent organic carbon", xlab = NULL)

```

#### Soil carbon density, Nakorn

```{r, echo = FALSE}

n_cdens_mc <- moran.mc(nakorn_sp$sbplt_c_, nbw, nsim=99)
n_cdens_mc

plot(n_cdens_mc, main = "Spatial autocorrelation: Nakorn soil C density", xlab = NULL)

```

## Quick mixed effects model of species effect

Filter out dominant species with n = 1 from the dataset, just for now.

```{r, echo = FALSE}

genus_cnts <- dat %>%
  drop_na() %>%
  group_by(site, dom_genus) %>%
  add_count(dom_genus) %>%
  select(site, plot, subplot, dom_genus, n) %>%
  distinct

mdl_dat <- dat %>%
  left_join(genus_cnts, by = c("site", "plot", "subplot")) %>%
  select(-dom_genus.y) %>%
  rename(dom_genus = dom_genus.x)

filt_mdl_dat <- mdl_dat %>%
  filter(n >= 3)

filt_mdl_dat %>%
  select(site, dom_genus, n) %>%
  distinct()

```

#### Looking at filtered data:

```{r, echo = FALSE}

filt_mdl_dat %>%
  ggplot() + 
  geom_point(aes(x = dom_genus, y = subplot_bd, col = site)) +
  xlab("Dominant genus") +
  ylab("Bulk density (g/cm^3)") +
  ggtitle("Bulk density  ~ dominant genus")

```

```{r, echo = FALSE}

filt_mdl_dat %>%
  ggplot() + 
  geom_point(aes(x = dom_genus, y = subplot_poc, col = site)) +
  xlab("Dominant genus") +
  ylab("Percent organic C (%)") +
  ggtitle("Percent organic C  ~ dominant genus")

```

Percent organic carbon ~ dominant genus:

```{r, echo = FALSE}

filt_mdl_dat %>%
  ggplot() + 
  geom_point(aes(x = dom_genus, y = subplot_c_dens, col = site)) +
  xlab("Dominant genus") +
  ylab("Soil C density (mg/cm^3)") +
  ggtitle("Soil C density  ~ dominant genus")

```


Subset to just Krabi, given that there is only n = 2 per transect in Nakorn and thus have singular convergence.

Run 2 models:

1. Null model with percent organic carbon predicted by total biomass, random intercept by plot (account for plotwise correlation)
2. LME model with percent organic carbon predicted by dom genus + total biomass, random intercept by plot (account for plotwise correlation)

Run an ANOVA between the two models to test for significance of alternative model.

#### Null model

```{r}

null <- lme4::lmer(subplot_poc ~ subplot_ttl_ha + (1|plot), 
                   data = filter(filt_mdl_dat, site == "Krabi"), REML = FALSE)
summary(null)

```

#### LME model

```{r}

lmm <- lme4::lmer(subplot_poc ~ subplot_ttl_ha + dom_genus + (1|plot), 
                  data = filter(filt_mdl_dat, site == "Krabi"), REML = FALSE)
summary(lmm)

```

#### Model comparison via ANOVA

```{r}

anova(null, lmm)

```

## Same comparison for bulk density

#### Null model

```{r}

null <- lme4::lmer(subplot_bd ~ subplot_ttl_ha + (1|plot), 
                   data = filter(filt_mdl_dat, site == "Krabi"), REML = FALSE)
summary(null)

```

#### LME model

```{r}

lmm <- lme4::lmer(subplot_bd ~ subplot_ttl_ha + dom_genus + (1|plot), 
                  data = filter(filt_mdl_dat, site == "Krabi"), REML = FALSE)
summary(lmm)

```

#### Model comparison via ANOVA

```{r}

anova(null, lmm)

```




## Same comparison for soil carbon density

#### Null model

```{r}

null <- lme4::lmer(subplot_c_dens ~ subplot_ttl_ha + (1|plot), 
                   data = filter(filt_mdl_dat, site == "Krabi"), REML = FALSE)
summary(null)

```

#### LME model

```{r}

lmm <- lme4::lmer(subplot_c_dens ~ subplot_ttl_ha + dom_genus + (1|plot), 
                  data = filter(filt_mdl_dat, site == "Krabi"), REML = FALSE)
summary(lmm)

```

#### Model comparison via ANOVA

```{r}

anova(null, lmm)

```


## Quick summary:

Find significant effect of species on percent organic carbon, but not bulk density or soil carbon density (product of POC and bulk density).

Tends to align with what we'd expect given species vs. environmental setting influences? bulk density determined by sedimentation, whereas organic matter input to soil can be influence by vegetation morphology.
